Probabilistic Catalogs for Crowded Stellar Fields

### 2013-04-08 paper/ms.tex
More verbose captions

For simplicity, we assume a uniform prior probability distribution for the
position of each star. The use of independent priors for the positions creates
a strong preference for catalogs where
the stars are uniformly scattered across the image. Thus, this model is
appropriate for small images where the density of stars is approximately
uniform across the imaeg. In other scenarios, such as images of stellar clusters, it is possible
to parameterize the spatial distribution of the stars in a similar way to how
we have parameterized the luminosity function, i.e. as a hierarchical model.

@@

\begin{table}[ht!]\footnotesize
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
Parameter & Value (Test Case 1) & Value (Test Case 2)\\
\hline
$N$ & 100 & 1000\\
$h_1$ & 0.3 & 0.3\\
$h_2$ & 0.6 & 0.6\\
$\alpha_1$ & 1.1 & 1.1\\
$\alpha_2$ & 2 & 2\\
\hline
$\sigma_0$ & 10 & 10\\
$\eta$ & 10 & 10 \\
$s_1$ & 0.02 & 0.02\\
$s_2$ & 0.1 & 0.1\\
$w$ & 0.5 & 0.5\\
\hline
\end{tabular}
\end{center}
\caption{True parameter values used to generate the simulated data. The only
difference between the two images is that Test Case 2 contains more stars than
Test Case 1.
\label{tab:truth}}
\end{table}

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=\textwidth]{Figures/test_cases.eps}
\caption{The two simulated images used to test our methodology.
{\bf Left}: A field containing $\sim$100 stars.
{\bf Right}: A field containing $\sim$1000 stars.\label{fig:simulated_data}}
\end{center}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=\textwidth]{Figures/inference1.eps}
\end{center}
\caption{Inference about the parameters for Test Case 1. Note that there is
considerable uncertainty (particularly about $h_2$), which occurs because the
shape of the broken power-law does not depend strongly on the parameters.
The true input values are plotted as filled squares.
\label{fig:results1}}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=\textwidth]{Figures/inference2.eps}
\end{center}
\caption{Inference about the parameters for Test Case 2. Note that the
parameters are still not very well constrained even with the larger number of
stars. This occurs because the fluxes of faint stars are not accurately measured
and because the shape of the broken power-law distribution does not vary rapidly
as a function of the parameters. The true input values are plotted as filled squares.\label{fig:results2}}
\end{figure}

\begin{figure}[ht!]
\begin{center}
\includegraphics[scale=0.6]{Figures/catalogs.eps}
\end{center}
\caption{Fundamentally, the output from our method is samples from the posterior
distribution over the catalog space. Nine example catalogs are shown, sampled from the posterior distribution for
Test Case 2. Features in common represent features with high probability,
and differences between the catalogs represent conclusions that are uncertain.
The area of each circle is proportional to the flux of the star.
The posterior samples may be used to compute summary images; these are presented
in Figure~\ref{fig:summaries}.
\label{fig:catalogs}}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=\textwidth]{Figures/summaries.eps}
\end{center}
\caption{Summary images for Test Case 2. The upper left panel shows the posterior
mean high-resolution scene. The upper right panel shows the posterior mean scene
when observed at the resolution of the data, and the bottom panels show the
standardized model residuals.
\label{fig:summaries}}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=\textwidth]{Figures/luminosity_function.eps}
\end{center}
\caption{The cumulative luminosity functions produced by the Bayesian method
(several posterior samples shown)
and
\sex~(for various values of the threshold parameters), compared with the
actual cumulative LF. Both methods correctly identify the fluxes at the bright
end, with some uncertainty due to overlapping sources. However, at the lower
end \sex~is unable to detect all of the stars whereas the true CLF is typical
of the posterior distribution.
\label{fig:luminosity_function}}
\end{figure}

>>>>

For simplicity, we assume a uniform prior probability distribution for the
position of each star. The use of independent priors for the positions creates
a strong preference for catalogs where
the stars are uniformly scattered across the image. Thus, this model is
appropriate for small images where the density of stars is approximately
uniform across the image. In other scenarios, such as images of stellar clusters, it is possible
to parameterize the spatial distribution of the stars in a similar way to how
we have parameterized the luminosity function, i.e. as a hierarchical model.

@@

\begin{table}[ht!]\footnotesize
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
Parameter & Value (Test Case 1) & Value (Test Case 2)\\
\hline
$N$ & 100 & 1000\\
$h_1$ & 0.3 & 0.3\\
$h_2$ & 0.6 & 0.6\\
$\alpha_1$ & 1.1 & 1.1\\
$\alpha_2$ & 2 & 2\\
\hline
$\sigma_0$ & 10 & 10\\
$\eta$ & 10 & 10 \\
$s_1$ & 0.02 & 0.02\\
$s_2$ & 0.1 & 0.1\\
$w$ & 0.5 & 0.5\\
\hline
\end{tabular}
\end{center}
\caption{True parameter values used to generate the simulated data. 
$N$ is the number of stars, $h_1$ and $h_2$ are the lower limit and break
point of the flux distribution respectively, and $\alpha_1$ and $\alpha_2$ 
are the slopes of the flux distribution. $\sigma_0$ and $\eta$ describe the
noise properties, and $s_1$, $s_2$, and $w$ are the PSF parameters. The only
difference between the two images is that Test Case 2 contains more stars than
Test Case 1.
\label{tab:truth}}
\end{table}

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=\textwidth]{Figures/test_cases.eps}
\caption{The two simulated images used to test our methodology.
{\bf Left}: An image containing $\sim$100 stars.
{\bf Right}: An image containing $\sim$1000 stars.\label{fig:simulated_data}}
\end{center}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=\textwidth]{Figures/inference1.eps}
\end{center}
\caption{Inference about the parameters for Test Case 1. The left panel shows
the posterior distribution for the number of stars $N$, and the right
panels show the joint posterior distributions for the flux distribution
(hyper)parameters. Note that there is
considerable uncertainty (particularly about $h_2$), which occurs because the
shape of the broken power-law does not depend strongly on the parameters.
The true input values are plotted as filled squares.
\label{fig:results1}}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=\textwidth]{Figures/inference2.eps}
\end{center}
\caption{Inference about the parameters for Test Case 2. Note that the
flux distribution (hyper)parameters are still not very well constrained even with
the larger number of
stars. This occurs because the fluxes of faint stars are not accurately measured
and because the shape of the broken power-law distribution does not vary rapidly
as a function of its parameters. The true input values are plotted as filled squares.\label{fig:results2}}
\end{figure}

\begin{figure}[ht!]
\begin{center}
\includegraphics[scale=0.6]{Figures/catalogs.eps}
\end{center}
\caption{Fundamentally, the output from our method is samples from the posterior
distribution over the catalog space. Nine example catalogs are shown, sampled from the posterior distribution for
Test Case 2. Features in common represent features with high probability,
and differences between the catalogs represent conclusions that are uncertain.
The area of each circle is proportional to the flux of the star.
The posterior samples may be used to compute summary images; some these are presented
in Figure~\ref{fig:summaries}.
\label{fig:catalogs}}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=\textwidth]{Figures/summaries.eps}
\end{center}
\caption{Summary images produced from the posterior distribution for Test Case 2. The upper left panel shows the posterior
mean high-resolution scene. The upper right panel shows the posterior mean scene
when observed at the resolution of the data, and the bottom panels show the
standardized model residuals.
\label{fig:summaries}}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=\textwidth]{Figures/luminosity_function.eps}
\end{center}
\caption{The cumulative luminosity functions (number of stars above a given
flux, as a function of flux) produced by the Bayesian method
(several posterior samples shown)
and
\sex~(for various values of the threshold parameters), compared with the
actual cumulative LF. Both methods correctly identify the fluxes at the bright
end, with some uncertainty due to overlapping sources. However, at the lower
end \sex~is unable to detect all of the stars whereas the true CLF is typical
of the posterior distribution.
\label{fig:luminosity_function}}
\end{figure}
